---
title: "SPOOKY Data Analysis"
author: "Mao Guan"
date: "January 30, 2018"
output:
  html_document:
    df_print: paged
---
# Introduction 
"As I scurried across the candlelit chamber, manuscripts in hand, I thought I'd made it. Nothing would be able to hurt me anymore. Little did I know there was one last fright lurking around the corner."
This is a part of a horror story which terrifies and excites us. In the passage, we are going to analyze the data composed of horror stories written by Edgar Allan Poe, Mary Shelley, and HP Lovecraft. The data was prepared by chunking larger texts into sentences using CoreNLP's MaxEnt sentence tokenizer. Specifically we would like to consider the similarities and the differences between the texts attributed to each author and study patterns that could be used to characterize the writing styles of the three authors.


#Libraries Preparation
```{r, message = F, warning = F}
packages.used=c("corrgram","RSentiment","xgboost","ggridges")

# check packages that need to be installed.
packages.needed=setdiff(packages.used, 
                        intersect(installed.packages()[,1], 
                                  packages.used))
# install additional packages
if(length(packages.needed)>0){
  install.packages(packages.needed, dependencies = TRUE)
}

# general visualisation
library('ggplot2') # visualisation
library('scales') # visualisation
library('grid') # visualisation
library('gridExtra') # visualisation
library('RColorBrewer') # visualisation
library('corrplot') # visualisation

# general data manipulation
library('dplyr') # data manipulation
library('readr') # input/output
library('data.table') # data manipulation
library('tibble') # data wrangling
library('tidyr') # data wrangling
library('stringr') # string manipulation
library('forcats') # factor manipulation

# specific visualisation
#library('alluvial') # visualisation
library('ggrepel') # visualisation
#library('ggfortify')# visualisation
library('ggridges') # visualisation
#library('gganimate') # visualisation
#library('ggExtra') # visualisation
library('corrgram')

# specific data manipulation
library('lazyeval') # data wrangling
library('broom') # data wrangling
library('purrr') # string manipulation
library('reshape2') # data wrangling

# Text / NLP
library('tidytext') # text analysis
library('tm') # text analysis
library('SnowballC') # text analysis
library('topicmodels') # text analysis
library('wordcloud') # test visualisation
library('igraph') # visualisation
#library('ggraph') # visualisation
#library('babynames') # names
library("RSentiment")

# Models
library('Matrix')
library('xgboost')
#library('caret')

#library('treemapify') #visualisation
source("../libs/multiplot.R")
```
#Data Preparations
First, before we dive into the data. Let's take a glimpse of the data offered.
```{r}
spookydata = read.csv('../data/spooky.csv', as.is = TRUE)
```
Then, let's find whether the sentence length among the authors will vary much.
```{r}
p <- spookydata %>%
  mutate(sen_len = str_length(text)) %>%
  ggplot(aes(sen_len, author, fill = author)) +
  geom_density_ridges() +
  scale_x_log10() +
  theme(legend.position = "right") +
  labs(x = "Sentence length")

plot(p)
```
Looks like the three authors' sentence length distribution varies. HP Lovecraft prefers long sentence and is more focused on using sentences with length around 200.


Second, let's do some simple treatment to our data: remove the invalid information incluing tokens.
```{r}
spooky_wrd <- spookydata %>%
   unnest_tokens(word, text) %>%
   # remove stopwords
   anti_join(stop_words, by = "word") %>%
   count(author, word) %>%
   ungroup()
```
##Word Clouds
In this part, lets' make a word cloud to see the most common words used by the three authors together and separately

```{r}
spooky_wrd_all <- spooky_wrd %>%
  group_by(word) %>%
  summarise(n = sum(n)) %>%
  ungroup()

wordcloud(spooky_wrd_all$word, spooky_wrd_all$n,
          max.words = 200, scale = c(2.0,0.5),
          colors = RColorBrewer::brewer.pal(9, "YlOrRd")[4:10])
```
Those words are the 80 most common words in the datasets, we have seen that the authors would like to use the word like "life","death","door" and "light". It seems that the horror fictions would like to decorate the normal life with life and death.
Naturally, we would assume some diffrence between different authors. Now, let's see if there is any difference between them.
```{r}

spooky_wrd_MWS <- spooky_wrd %>%
  filter(author == "MWS") %>%
  group_by(word) %>%
  summarise(n = sum(n)) %>%
  ungroup()
wordcloud(spooky_wrd_MWS$word, spooky_wrd_MWS$n,
          max.words = 200, scale = c(2.0,0.5),
          colors = RColorBrewer::brewer.pal(9, "YlOrRd")[4:10])

```
It seems that Mary Shelley is more focused on the human body and relationships.

```{r}
spooky_wrd_EAP <- spooky_wrd %>%
  filter(author == "EAP") %>%
  group_by(word) %>%
  summarise(n = sum(n)) %>%
  ungroup()
wordcloud(spooky_wrd_MWS$word, spooky_wrd_MWS$n,
          max.words = 200, scale = c(2.0,0.5),
          colors = RColorBrewer::brewer.pal(9, "YlOrRd")[4:10])
```

As for Edgar Allan Poe, it seems that he is more concerned about the time. Maybe he is the kind of writer who is good at creating the atmosphere of emergencies.
```{r}
spooky_wrd_HPL <- spooky_wrd %>%
  filter(author == "HPL") %>%
  group_by(word) %>%
  summarise(n = sum(n)) %>%
  ungroup()
wordcloud(spooky_wrd_MWS$word, spooky_wrd_MWS$n,
          max.words = 200, scale = c(2.0,0.5),
          colors = RColorBrewer::brewer.pal(9, "YlOrRd")[4:10])
```
Well, HP Lovecraft is different from the other two that 
he prefers the night. Perhaps, his story happens mostly at night.



#Similarity of authors
```{r}
simi_author <- spooky_wrd %>%
   spread(author, n, fill = 0) %>%
   na.omit()

simi_word <- spooky_wrd %>%
   spread(word, n, fill = 2) %>%
   na.omit()

corrgram(simi_author, lower.panel = panel.shade, upper.panel = panel.cor)

```




Take Edgar Allan Poe as consideration, comparing to HP Lovecraft and Mary Shelley. We witnessed that there are certain words like "found", "hand" and etc. are in common.

It will also be interesting to see whether there is any major difference among the authors about the genders emphasis of the fictions.

```{r}
p  <- spookydata %>%
  unnest_tokens(word, text) %>%
  
  mutate(female = ( word == "she" | word == "her" | word == "hers" | word == "female" | word == "girl"|
                    word == "woman" | word == "lady" | word == "madam" |
                    word == "women" )) %>%
  mutate(male = ( word == "he" | word == "him" | word == "his" | word == "male" |word == "boy"|
                    word == "man" | word == "gentleman" | word == "sir" |
                    word == "lord" | word == "men" )) %>%
  unite(sex, female, male) %>%
  mutate(sex = fct_recode(as.factor(sex), male = "FALSE_TRUE", 
                          female = "TRUE_FALSE", other = "FALSE_FALSE")) %>%
  filter(sex != "other") %>%
  ggplot(aes(sex, fill = author)) +
  labs(x = "Gender Difference") +
  geom_bar(position = "dodge")

plot(p)
```
Apparently, male characters are the leading gender in the horror fictions. Separately, HP Lovecraft is much more focused on the male characters than the other authors. Then, what exactly words did they use as gender indications. Let's find it out.
```{r}
p1 <- spookydata %>%
  unnest_tokens(word, text) %>%
  filter((word == "him") | (word == "his") | (word == "man") | (word == "gentleman") | (word == "male") | (word == "lord") |(word == "boy")| (word == "he") | (word == "men")) %>%
  ggplot(aes(word, fill = author)) +
  geom_bar(position = "dodge")

p2 <- spookydata %>%
  unnest_tokens(word, text) %>%
  filter(( word == "she" | word == "her" | word == "hers" | word == "female" | word == "girl"|
                    word == "woman" | word == "lady" | word == "madam" |
                    word == "women" )) %>%
  ggplot(aes(word, fill = author)) +
  geom_bar(position = "dodge")

layout <- matrix(c(1,2),2,1,byrow=TRUE)
multiplot(p1, p2, layout=layout)

```

Apparently, the simple words"he, him, his" and "her","she" are the most common words in the charts. And it is of no apparent difference among the authors about the preference of the words in each gender.

```{r}
t_heshe <- spookydata %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
  separate(bigram, c("word1", "genderword"), sep = " ")

t_heshe %>%
  filter(word1 == "she" | word1 == "he" ) %>%
  group_by(author, word1, genderword) %>%
  count() %>%
  ungroup() %>%
  group_by(author, word1) %>%
  top_n(5,n) %>%
  ggplot(aes(genderword, n, fill = author)) +
  geom_col() +
  scale_y_log10() +
  coord_flip() +
  facet_grid(word1 ~ author)

t_himh <- spookydata %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2) %>%
  separate(bigram, c("word1", "genderword"), sep = " ")

t_himh %>%
  filter(word1 == "her" | word1 == "him" ) %>%
  group_by(author, word1, genderword) %>%
  count() %>%
  ungroup() %>%
  group_by(author, word1) %>%
  top_n(5,n) %>%
  ggplot(aes(genderword, n, fill = author)) +
  geom_col() +
  scale_y_log10() +
  coord_flip() +
  facet_grid(word1 ~ author)

```
It is easy to see that Lovecraft is apparently less interested in using "she" and "her" than the other authors. And EAP and MWS are quite alike in using words associated with "she","he","her"and "him". We may infer that Lovecraft' writing styles in characters are different from the other two while the other two are quite alike.

Okay, after analysis of the gender difference among the authors, why don't we take a look at the dialogs of the fictions. It is know that the dialogs will play a key role in displaying the horrified atmosphere out.Then how is the dialogs differering among the authors?







#Sentiment Analysis Part
##Most frequent words
###tf-idf
For the speeches, we remove extra white space, convert all letters to the lower case, remove stop words, removed empty words due to formatting errors, and remove punctuation.

After cleaning up corpus, we compute TDM. Documemnt Term Matrix (DTM) or Term Document Matrix (TDM) is a document that lists all occurrences of words of each document.
```{r}
corpus_eap <-  Corpus(VectorSource(spooky_wrd_EAP$word))
corpus_eap<-tm_map(corpus_eap, content_transformer(tolower))
corpus_eap<-tm_map(corpus_eap, removeWords, stopwords("english"))
corpus_eap<-tm_map(corpus_eap, removeWords, character(0))
corpus_eap<-tm_map(corpus_eap, removePunctuation)
tdm.all_eap<-TermDocumentMatrix(corpus_eap)
tdm.tidy_eap=tidy(tdm.all_eap)
tdm.overall_eap=summarise(group_by(tdm.tidy_eap, term), sum(count))

corpus_hpl <-  Corpus(VectorSource(spooky_wrd_HPL$word))
corpus_hpl<-tm_map(corpus_hpl, content_transformer(tolower))
corpus_hpl<-tm_map(corpus_hpl, removeWords, stopwords("english"))
corpus_hpl<-tm_map(corpus_hpl, removeWords, character(0))
corpus_hpl<-tm_map(corpus_hpl, removePunctuation)
tdm.all_hpl<-TermDocumentMatrix(corpus_hpl)
tdm.tidy_hpl=tidy(tdm.all_hpl)
tdm.overall_hpl=summarise(group_by(tdm.tidy_hpl, term), sum(count))

corpus_mws <-  Corpus(VectorSource(spooky_wrd_MWS$word))
corpus_mws<-tm_map(corpus_mws, content_transformer(tolower))
corpus_mws<-tm_map(corpus_mws, removeWords, stopwords("english"))
corpus_mws<-tm_map(corpus_mws, removeWords, character(0))
corpus_mws<-tm_map(corpus_mws, removePunctuation)
tdm.all_mws<-TermDocumentMatrix(corpus_mws)
tdm.tidy_mws=tidy(tdm.all_mws)
tdm.overall_mws=summarise(group_by(tdm.tidy_mws, term), sum(count))


```

##Emotion



```{r}
spooky_wrd <- spookydata %>% unnest_tokens(word, text)%>%
  anti_join(stop_words, by = "word")
```

```{r}
spooky_wrd %>%
    filter(author == "MWS") %>%
  inner_join(get_sentiments("loughran"), by = "word") %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup() %>%
  group_by(sentiment) %>%
  top_n(8, n) %>%
  mutate(word = reorder(word, n)) %>%
  ungroup() %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  coord_flip()+
  labs( x = NULL,y = "Sentiment Analysis") +
  ggtitle("Mary Shelley: Negative  Positive Words")
```


```{r}
spooky_wrd %>%
    filter(author == "HPL") %>%
  inner_join(get_sentiments("loughran"), by = "word") %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup() %>%
  group_by(sentiment) %>%
  top_n(8, n) %>%
  mutate(word = reorder(word, n)) %>%
  ungroup() %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  coord_flip()+
  labs( x = NULL,y = "Sentiment Analysis") +
  ggtitle("H P Lovecraft: Negative  Positive Words")
```






```{r}
spooky_wrd %>%
    filter(author == "EAP") %>%
  inner_join(get_sentiments("loughran"), by = "word") %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup() %>%
  group_by(sentiment) %>%
  top_n(8, n) %>%
  mutate(word = reorder(word, n)) %>%
  ungroup() %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  coord_flip()+
  labs( x = NULL,y = "Sentiment Analysis") +
  ggtitle("E A Poe: Negative  Positive Words")
```
```{r}
spooky_wrd %>%
  inner_join(get_sentiments("loughran"), by = "word") %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("#F8766D", "#00BFC4"), max.words = 500)
```
#The super negative Index

Define the super negative Index = (#uncertainty+#Negative + #constraining)/(#Postive + #Negative + #constraining + #uncertainty)

```{r}
pic1 <- spooky_wrd %>%
  inner_join(get_sentiments("loughran"), by = "word") %>%
  ggplot(aes(author, fill = sentiment)) +
  geom_bar(position = "fill")

pic2 <- spooky_wrd %>%
  inner_join(get_sentiments("loughran"), by = "word") %>%
  group_by(author, id, sentiment) %>%
  count() %>%
  spread(sentiment, n, fill = 0) %>%
  group_by(author, id) %>%
  summarise(neg = sum(negative),
            con = sum(constraining),
            unc = sum(uncertainty),
            pos = sum(positive)) %>%
  arrange(id) %>%
  mutate(frac_neg = 1 - pos/(pos + neg + con+unc)) %>%
  ggplot(aes(frac_neg, fill = author)) +
  geom_density(bw = .3, alpha = 0.5) +
  theme(legend.position = "right") +
  labs(x = "d")

layout <- matrix(c(1,2),1,2,byrow=TRUE)
multiplot(pic1, pic2, layout=layout)

```
#Topic Models
First, let's explore the structure using Latent Dirichlet Allocation Moddel(LDA). This method yields an unsupervised classifictaion of documents. 
By seeking the clusters corresponding to different topics, we will be able to find underlting structure in the data.

To analyze it, we need first transform the data into the document term matrices(DTM). 
```{r}
spooky_wrd<-spookydata %>% unnest_tokens(word, text)

freq <- spooky_wrd %>%
  anti_join(stop_words, by = "word") %>%
        count(id, word)
spooky_tm <- cast_dtm(freq, id, word, n)
spooky_tm
```
Now, let's perform LDA for 10 topics. We may assume that each of our authors contribute to more than one story to our text.

```{r}
spooky_lda <- LDA(spooky_tm, k = 10, method = "Gibbs", control = list(seed = 1234))
```
After training, let's visualize the final effects of the data.


```{r}

spooky_lda.topics <- as.matrix(topics(spooky_lda))
table(c(1:10, spooky_lda.topics))
write.csv(spooky_lda.topics, file = paste("../out/LDA_Gibbs",10,"Doc_Topics.csv"))
spooky_topic <- spooky_lda.topics
```
As the chart show, beta indicates the probability that this particular word belongs to that topic.
```{r}
spooky_lda.terms <- as.matrix(terms(spooky_lda,5))
spooky_lda.terms
```


```{r}
chapter_topics <- tidy(spooky_lda, matrix = "beta")

 chapter_topics%>%
  group_by(topic) %>%
  top_n(5, beta) %>%
  ungroup() %>%
  arrange(topic, -beta) %>%
  mutate(topics = reorder(term, beta)) %>%
  ggplot(aes(topics, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free", ncol = 5) +
  coord_flip()
```

We may wonder how are the topics assocaited with each author?
```{r}
eap <- spooky_wrd %>%
   filter(author == 'EAP')
hpl <- spooky_wrd %>%
   filter(author == 'HPL')
mws <- spooky_wrd %>%
   filter(author == 'MWS')


```

```{r}
freq <- eap %>%
    anti_join(stop_words, by = "word") %>%
        count(id, word)
spooky_tm_EAP <- cast_dtm(freq, id, word, n)

freq <- hpl %>%
    anti_join(stop_words, by = "word") %>%
        count(id, word)
spooky_tm_HPL <- cast_dtm(freq, id, word, n)

freq <- mws %>%
    anti_join(stop_words, by = "word") %>%
        count(id, word)
spooky_tm_MWS <- cast_dtm(freq, id, word, n)

```



```{r}
spooky_lda_EAP <- LDA(spooky_tm_EAP, method = "Gibbs", k = 10, control = list(seed = 1234))
spooky_lda_HPL <- LDA(spooky_tm_HPL, method = "Gibbs", k = 10, control = list(seed = 1234))
spooky_lda_MWS <- LDA(spooky_tm_MWS, method = "Gibbs", k = 10, control = list(seed = 1234))

```

```{r}
spooky_lda_EAP.topics <- as.matrix(topics(spooky_lda_EAP))
spooky_lda_EAP.terms <- as.matrix(terms(spooky_lda_EAP,5))
spooky_lda_HPL.topics <- as.matrix(topics(spooky_lda_HPL))
spooky_lda_HPL.terms <- as.matrix(terms(spooky_lda_HPL,5))
spooky_lda_MWS.topics <- as.matrix(topics(spooky_lda_MWS))
spooky_lda_MWS.terms <- as.matrix(terms(spooky_lda_MWS,5))
```

```{r}
spooky_topic_EAP <- tidy(spooky_lda_EAP, matrix = "beta")
spooky_topic_EAP %>% sample_n(5)
spooky_topic_HPL <- tidy(spooky_lda_HPL, matrix = "beta")
spooky_topic_HPL %>% sample_n(5)
spooky_topic_MWS <- tidy(spooky_lda_MWS, matrix = "beta")
spooky_topic_MWS %>% sample_n(5)
```

```{r}
spooky_topic_EAP %>%
  group_by(topic) %>%
  top_n(5, beta) %>%
  ungroup() %>%
  arrange(topic, -beta) %>%
  mutate(topics = reorder(term, beta)) %>%
  ggplot(aes(topics, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free", ncol = 5) +
  coord_flip()+ggtitle("E A Poe topic modelling")
```
It is not hard to see that E.A.Poe's ten topics distribution.Within topic1, he likes to quote the words:life, fell, air, moment, night and expression. 

```{r}
spooky_topic_HPL %>%
  group_by(topic) %>%
  top_n(5, beta) %>%
  ungroup() %>%
  arrange(topic, -beta) %>%
  mutate(topics = reorder(term, beta)) %>%
  ggplot(aes(topics, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free", ncol = 5) +
  coord_flip()+ggtitle("H P Lovecraft topic modelling")
```
Different from E.A.Poe, H.P.Lovecraft's focus on topic 1 lies on the words:black, dr, time, half and wild.
```{r}
spooky_topic_MWS %>%
  group_by(topic) %>%
  top_n(5, beta) %>%
  ungroup() %>%
  arrange(topic, -beta) %>%
  mutate(topics = reorder(term, beta)) %>%
  ggplot(aes(topics, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free", ncol = 5) +
  coord_flip()+ggtitle("Mary Shelley topic modelling")
```
What's unique about Mary Shelley is that she focuses on the words: father, life, power, friend, passed on topic 1. In all, by topic modelling, we shall see the differences between the authors. For example, with the word "friend", Mary Shelley would like to write friend together with the words: life, power, father and passed. H P Lovecraft would like to write friend together with the words:fear, west, floor, held. E A Poe would like to write friend together with the words:found, body, house, madame. We may observe this difference in the way like this, and this is the difference of the writing styles among the authors.
















